Peer-to-Peer Permissionless Consensus via Reputation

### 2023-06-08 freechains/freechains.tex
...

Content publishing in Internet forums and social media is
increasingly more centralized in a few companies (e.g. Facebook and
Twitter)~\cite{internet.fixing,p2p.osn,p2p.dosn}.
%These companies benefit from closed protocols and network effects to keep
%users locked in their platforms.
On the one hand, these companies offer free storage, friendly user interfaces,
and robust access.
On the other hand, they concentrate too much power over its users, holding and
monetizing over their private data and behavior.
%
Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
strive to achieve consistency when dealing with malicious users.
%given the decentralization of authority and infrastructure.
% and push to end users the responsibility to manage data and connectivity.

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In a centralized system, items (i) and (ii) are trivially achieved assuming
availability and delivery order in the service, while for item (iii), users
have to trust the service to moderate content.
In a decentralized setting, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively replicate and
disseminate posts among peers until they reach all
users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters and blacklists, but which only apply a posteriori, when the
abusive messages have already been flooded in the network.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybil attacks~\cite{p2p.sybil}.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech).
For instance, Bitcoin transactions permanently store arbitrary human-authored
content~\cite{btc.data}, including harmful texts and images~\cite{btc.content}.

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
scarce resources themselves, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow users to create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer content dissemination protocol that provides strong
eventual consistency~\cite{p2p.crdts,p2p.sec}.
%
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

The rest of the paper is organized as follows:
In Section~\ref{sec.design}, we describe the design of the reputation and
consensus mechanism for public forums and discuss some security threats.
In Section~\ref{sec.freechains}, we describe the concrete reputation rules we
implemented for \FC and evaluate the performance of the protocol in real-world
public forums.
In Section~\ref{sec.related}, we compare our system with publish-subscribe
protocols, federated applications, and fully peer-to-peer systems.
In Section~\ref{sec.conclusion}, we conclude this work.

@@

In the absence of moderation, permissionless peer-to-peer forums are
impractical, mostly because of Sybils abusing the system.
For instance, it should take a few seconds to generate thousands of fake
identities and SPAM millions of messages into the system.
%Hence, without moderation, there are no limits on the number and size of posts
%and no reasonable policy to distinguish quality.
For this reason, we propose a reputation system that works together with a
consensus algorithm to resist Sybil attacks.
% and make peer-to-peer public forums practical.

\begin{table}
\centering
\includegraphics[width=0.75\textwidth]{general.png}
\caption{General reputation rules in public forums.}
\label{fig.general}
\end{table}

@@

Our key idea to stablish consensus in forum DAGs is to favor forks with posts
from users that constitute the majority of the reputation in the network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
%
We assume that within the common prefix, users $a$ and $b$ have contributed
with better content and therefore have more reputation combined than $c$ has
alone (i.e., $8+5>3$).
We will use this information when ordering branches in the consensus algorithm.
%
Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
new users $x$ and $y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to common past posts could
possibly block posts from users in the other branch, depending on how they are
ordered.
Therefore, we need to order the branches with a global consensus, such that we
can validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

Therefore, based on the reputation of users in the common prefix,
Figure~\ref{fig.reps}.B indicates consensus order between posts in the forum
DAG:
    the common prefix comes first,
    followed by all posts in \code{branch-2},
    followed by all posts in \code{branch-1},
    followed by the common suffix.
%
The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

@@

There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation need to reorder all
posts starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on reputation of an agreed common prefix,
    (iv) leading to a global and deterministic result.
%
Note that the consensus order exists for reputation accountability purposes,
and is only a view of the primary DAG structure.

@@

Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
intentionally disconnects from the network for weeks to cultivate fake
identities $x$ and $y$ and accumulate \reps.
When merging, these fake users become legitimate and now accumulate the
majority of \reps in the forum.
%
However, this kind of attack is refutable because, since the common prefix is
the basis of consensus, users in the branch with more reputation can still
react even a posteriori.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ in \code{branch-2} so that
a further merge invalidates all posts from \code{branch-1}, removing them from
the DAG.
%
This kind of attack is innocuous because it requires at least half of the \reps
accumulated in the common prefix.
%
The same countermeasure can be used against a minority of malicious users
colluding to boost or protect their own posts.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{n100-d7.png}
\caption{
    Stable consensus permanently freezes the order of posts once they cross
    the threshold.
    The unstable order may still be affected by incoming branches.
}
\label{fig.hard}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{state-revoked.png}
\caption{
    State machine of posts:
    \code{BLOCKED} posts are not linked in the DAG.
    \code{ACCEPTED} posts are linked and retransmitted.
    The payload of \code{REVOKED} posts are not retransmitted.
}
\label{fig.state}
\end{figure}

@@

Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary but public bag and
retransmit them for some visibility in the network.

@@

\FC~\cite{fcs.sbseg20} is an unstructured peer-to-peer topic-based
publish-subscribe protocol, in which each \emph{chain} is a replicated
Merkle-DAG representing the causal relationships between the messages.
%
The protocol operation is typical of publish-subscribe systems: an author
publishes a post to a chain, and subscribed users eventually receive the
message.
% by synchronizing DAGs.

@@

The most basic concern in public forums is to resist Sybils abusing the
chains.
Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
due to the lack of a central authority.
Viable alternatives include (i) building social trust graphs, in which users
already in the community vouch for new users, or (ii) imposing explicit costs
for new posts, such as proof of work.
%
We propose a mix between trust graphs and economic costs.
%
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{forum.png}
\caption{
    The \code{like} approves the newbie message into the \code{\#forum} DAG.
}
\label{fig.forum}
\end{figure}

@@

Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes
in the chains:
    (i) welcoming new users,
    (ii) measuring the quality of posts, and
    (iii) revoking abusive posts (SPAM, fake news, illicit content, etc).
%
The quality of posts is subjective and is up to users to judge them with likes,
dislikes, or simply abstaining.
%
This way, access to chains is permissionless in the sense that the actual peers
and identities behind posts are not directly assessed by the protocol, but
instead by the judgement of other users in the system.
%
The reputation of a given post is the difference between its likes and
dislikes, which can be used in end-user software for filtering and highlighting
purposes.
%
On the one hand, since \reps are finite, users need to ponder to avoid
indiscriminate expenditure.
On the other hand, since \reps are limited to at most \nreps{30} per author
(rule~\code{4.b}), users also have incentives to rate content.
Hence, these upper and lower limits work together towards the quality of the
chains.
%
Note that a dislike shrinks the chain economy since it removes \reps from both
the origin and target.
Finally, the actual contents of a post may be revoked if it has at least 3
dislikes, and more dislikes than likes (rule~\code{3}).
However, considering that \reps are scarce, dislikes are encouraged to combat
abusive behavior, but not to eliminate divergences of opinion.

@@

In this section, we perform experiments to evaluate the performance and
practicability of \FC and its consensus algorithm.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.

@@

For the newsgroup, we use \texttt{N=15} and \texttt{M=5}, which represents a
larger number of peers with few interconnections to stress the local-first
nature of the protocol.
For the chat, we use \texttt{N=5} and \texttt{M=3}, which represents a smaller
number of peers with more interconnections.
%
We executed each simulation $4$ times in a conventional desktop PC (\texttt{i7}
CPU, \texttt{8GB} RAM, \texttt{512GB} SSD).
Since the variations were negligible, we always discuss the median measures.

@@

Many other systems have been proposed for decentralized content
sharing~\cite{p2p.survey,p2p.ecosystem}.
Here we consider the classes of publish-subscribe, federated, and peer-to-peer
protocols.

@@

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
peer-to-peer gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.stan,pubsub.vitis,pubsub.gossipsub,pubsub.rappel}.
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies posts at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.

@@

Federated protocols, such as e-mail, allow users from one domain to seamlessly
exchange messages with users of other domains.
\emph{Diaspora}, \emph{Matrix}, and \emph{Mastodon} are recent federations for
social media, chat, and microblogging~\cite{p2p.ecosystem}, respectively.
%
As a drawback, identities in federations are not portable across domains, which
may become a problem when servers shutdown or users become unsatisfied with the
service~\cite{fed.distributed}.
In any of these cases, users have to grab their content, move to another
server, and announce a new identity to followers.
%
Moderation is also a major concern in federations~\cite{p2p.ecosystem}.
As an example, messages crossing domain boundaries may be subject to different
policies that might affect delivery.
With no coordinated consensus, it is difficult to make pervasive public forums
practical.
%
For this reason, Matrix supports a permissioned moderation system%
\footnote{Matrix moderation: \url{https://matrix.org/docs/guides/moderation}},
but which applies only within clients, after the messages have already been
flooded in the network.
%
As a counterpoint, federated protocols seem to be more appropriate for
real-time applications such as large chats rooms.
The number of hops and header overhead can be much smaller in client-server
architectures compared to peer-to-peer systems, which typically include message
signing, hash linking, and extra verification rules (as evaluated in
Section~\ref{sec.evaluation}).

@@

Regarding peer-to-peer protocols, Bitcoin~\cite{p2p.bitcoin} is probably the
most successful permissionless network, but serves specifically for electronic
cash.
IPFS~\cite{p2p.ipfs} and Dat~\cite{p2p.dat} are data-centric protocols for
hosting large files and applications, respectively.
Scuttlebutt~\cite{p2p.scuttlebutt} and Aether~\cite{p2p.ecosystem} are closer
to \FC goals and cover human communication.

>>>>

Content publishing in Internet forums and social media is
increasingly more centralized in a few companies (e.g. Facebook and
Twitter)~\cite{internet.fixing}.
%,p2p.osn,p2p.dosn}.
%These companies benefit from closed protocols and network effects to keep
%users locked in their platforms.
On the one hand, these companies offer free storage, friendly user interfaces,
and robust access.
On the other hand, they concentrate too much power over its users, holding and
monetizing over their private data and behavior.
%
Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
strive to achieve consistency when dealing with malicious users.
%given the decentralization of authority and infrastructure.
% and push to end users the responsibility to manage data and connectivity.

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In centralized systems, items (i) and (ii) are trivially achieved assuming
availability and delivery order, while item (iii) assumes that users trust the
service to moderate content.
In decentralized systems, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively disseminate posts
among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention}.
%,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to distinguish malicious behavior \emph{at the protocol
level}, and therefore, satisfy item (iii).
Without consensus, each client must rely on local anti-abuse policies, such as
SPAM filters and blacklists, but which only apply a posteriori, when the
abusive messages have already been flooded in the network.
%
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illicit content).
%At the very core, consensus is

@@

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybil attacks~\cite{p2p.sybil}.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects;
    (iii) they impose an external economic cost to use the protocol; and
    (iv)  they rely exclusively on objective rules to reach consensus.
%
These issues threaten our original decentralization goals, and more
importantly, they renounce subjectivity, which is inherent to content
moderation.
%
For instance, a unique timeline implies that all Internet content should be
subject to the same consensus rules, while objective rules alone cannot
distinguish abuse amid content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illicit content (e.g., hate speech).
For instance, Bitcoin transactions permanently store arbitrary human-authored
content, including harmful texts and images~\cite{btc.content}.
%~\cite{btc.data}

In this work, we adapt the idea of scarce resources to reach consensus and
eradicate Sybils, but in the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
scarce resources themselves, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are further used to evaluate other posts with
likes and dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization and network partitions, posts in a timeline form a
causal graph with only partial order, which we promote to a total order based
on the reputation of authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow users to create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated blockchain metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer (P2P) content dissemination protocol that provides
strong eventual consistency~\cite{p2p.sec}.
%p2p.crdts,
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

The rest of the paper is organized as follows:
In Section~\ref{sec.design}, we describe the design of the reputation and
consensus mechanism for public forums and discuss some security threats.
In Section~\ref{sec.freechains}, we describe the concrete reputation rules we
implemented for \FC and evaluate the performance of the protocol in real-world
public forums.
In Section~\ref{sec.related}, we compare our system with publish-subscribe
protocols, federated applications, and fully P2P systems.
In Section~\ref{sec.conclusion}, we conclude this work.

@@

In the absence of moderation, permissionless P2P forums are impractical, mostly
because of Sybils abusing the system.
For instance, it should take a few seconds to generate thousands of fake
identities and SPAM millions of messages.
%Hence, without moderation, there are no limits on the number and size of posts
%and no reasonable policy to distinguish quality.
For this reason, we propose a reputation system that works together with a
consensus algorithm to resist Sybil attacks.
% and make peer-to-peer public forums practical.

\begin{table}
\centering
\includegraphics[width=0.65\textwidth]{general.png}
\caption{General reputation rules in public forums.}
\label{fig.general}
\end{table}

@@

Our key idea to stablish consensus in forum DAGs is to favor forks with posts
from users that constitute the majority of the reputation.
These forks have more associated work and are analogous to longest chains in
Bitcoin.
%
In technical terms, we adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
presence of a network partition:
A common prefix has signed posts from users $a$, $b$, and $c$, when they were
still all connected.
%
We assume that within the prefix, users $a$ and $b$ have contributed with
better content and therefore have more reputation combined than $c$ has alone
(i.e., $8+5>3$).
We will use this fact when ordering branches in the consensus algorithm.
%
Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
$y$ with posts in \code{branch-1}.
In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
merge the branches with a common suffix.

Note that even non-malicious network partitions may experience conflicts when
merging branches.
For instance, dislike operations in one branch to common past posts could
possibly block posts from users in the other branch, depending on how they are
finally ordered.
Therefore, we need to order branches with a global consensus, such that we can
validate operations consistently across the network.
%
Note also that, to reach consensus, we should preferably rely on the maximum
information that all branches agree on, which is exactly the posts in the
common prefix.

Therefore, based on the reputation of users in the common prefix,
Figure~\ref{fig.reps}.B indicates the consensus order between posts in the forum
DAG:
    the common prefix comes first,
    then all posts in \code{branch-2},
    then all posts in \code{branch-1},
    then the common suffix.
%
The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
with user $c$ because, before the forking point, $a$ and $b$ have more
reputation than $c$, $x$, and $y$ combined.

@@

There are some other relevant considerations about forks and merges:
%
Peers that first receive branches with less reputation need to reorder all
posts after the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when peers
detect new longest chains. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to
append the other branch, with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the connected network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of reordering when
rejoining.

In summary, it is important to remark that
    (i) branches always introduce conflicts, which
    (ii) require consensus to order operations, which
    (iii) is based on a common prefix reputation,
    (iv) leading to a global and deterministic result.
%
Note that the consensus order exists only to account reputation and verify
operations, and is only a view of the primary DAG structure.

@@

Consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
intentionally disconnects from the network for weeks to cultivate fake
identities $x$ and $y$ with the intention to accumulate \reps.
When merging, these fake users would become legitimate and accumulate the
majority of \reps in the forum.
%
However, this kind of attack is refutable because, since the common prefix is
the basis of consensus, users in the branch with more reputation can still
react even a posteriori.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ so that a further merge
invalidates all posts from \code{branch-1}, removing them from the DAG.
%
This kind of attack is innocuous because it requires at least half of the \reps
accumulated in the common prefix.
%
The same countermeasure can be used against a minority of malicious users
colluding to boost or protect their own posts.

\begin{figure}
\centering
\includegraphics[width=0.85\textwidth]{n100-d7.png}
\caption{
    Stable consensus freezes the order of posts once they cross the threshold.
    The unstable order may still be affected by incoming branches.
}
\label{fig.hard}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{state-revoked.png}
\caption{
    State machine of posts:
    \code{BLOCKED} posts are not linked in the DAG.
    \code{ACCEPTED} posts are linked and retransmitted.
    The payload of \code{REVOKED} posts are not retransmitted.
}
\label{fig.state}
\end{figure}

@@

Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts nor retransmit them
to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
Therefore, a reasonable policy is to hold blocked posts in a temporary but
public bag and retransmit them for some visibility.

@@

\FC~\cite{fcs.sbseg20} is an unstructured P2P topic-based publish-subscribe
protocol, in which each \emph{chain} is a replicated Merkle-DAG representing
the causal relationships between the messages.
%
The protocol operation is typical of publish-subscribe systems: an author
publishes a post to a chain, and subscribed users eventually receive the
message.
% by synchronizing DAGs.

@@

The most basic concern in public forums is to resist Sybil attacks.
Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
lack of a central authority.
Viable alternatives include (i) building social trust graphs, in which users
already in the community vouch for new users, or (ii) imposing explicit costs
for new posts, such as proof of work.
%
We propose a mix between trust graphs and economic costs.
%
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

@@

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{forum.png}
\caption{
    The \code{like} approves the newbie message into the \code{\#forum} DAG.
}
\label{fig.forum}
\end{figure}

@@

Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes:
    (i) welcoming new users,
    (ii) measuring the quality of posts, and
    (iii) revoking abusive posts (SPAM, fake news, etc).
%
The quality of posts is subjective and is up to users to judge them with likes,
dislikes, or simply abstaining.
%
This way, access to chains is permissionless in the sense that the actual peers
and identities behind posts are not directly assessed by the protocol, but
instead by the judgement of other users in the system.
%
The reputation of a given post is the difference between its likes and
dislikes, which can be used in end-user software for filtering and highlighting
purposes.
%
On the one hand, since \reps are finite, users need to ponder to avoid
indiscriminate expenditure.
On the other hand, since \reps are limited to at most \nreps{30} per author
(rule~\code{4.b}), users also have incentives to rate content.
Hence, these upper and lower limits work together towards the quality of the
chains.
%
Note that a dislike shrinks the chain economy since it removes \reps from both
the origin and target.
Finally, the actual contents of a post may be revoked if it has at least 3
dislikes, and more dislikes than likes (rule~\code{3}).
However, considering that \reps are scarce, dislikes are encouraged to combat
abusive behavior, but not to eliminate divergences of opinion.

@@

We performed experiments to evaluate the performance of \FC and its consensus
algorithm.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.

@@

For the newsgroup, we use \texttt{N=15} and \texttt{M=5}, which represents a
larger number of peers with few interconnections to stress the local-first
nature of the protocol.
For the chat, we use \texttt{N=5} and \texttt{M=3}, which is a smaller number
of peers but with more interconnections.
%
We executed each simulation $4$ times in a desktop PC (\texttt{i7} CPU,
\texttt{8GB} RAM, \texttt{512GB} SSD).
Since the variations were negligible, we always discuss the median measures.

@@

Many other systems have been proposed for decentralized content
sharing~\cite{p2p.survey,p2p.ecosystem}.
Here we consider the classes of publish-subscribe, federated, and P2P
protocols.

@@

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
P2P gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.gossipsub}.
%,pubsub.stan,pubsub.vitis
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies posts at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.

@@

Federated protocols, such as e-mail, allow users from one domain to exchange
messages with users of other domains.
\emph{Diaspora}, \emph{Matrix}, and \emph{Mastodon} are recent federations for
social media, chat, and microblogging~\cite{p2p.ecosystem}, respectively.
%
As a drawback, identities in federations are not portable across domains, which
may become a problem when servers shutdown or users become unsatisfied with the
service~\cite{fed.distributed}.
In any of these cases, users have to grab their content, move to another
server, and announce a new identity to followers.
%
Moderation is also a major concern in federations~\cite{p2p.ecosystem}.
As an example, messages crossing domain boundaries may be subject to different
policies that might affect delivery.
With no coordinated consensus, it is difficult to make pervasive public forums
practical.
%
For this reason, Matrix supports a permissioned moderation system%
\footnote{Matrix moderation: \url{https://matrix.org/docs/guides/moderation}},
but which applies only within clients, after the messages have already been
flooded in the network.
%
As a counterpoint, federated protocols seem to be more appropriate for
real-time applications such as large chats rooms.
The number of hops and header overhead can be much smaller in client-server
architectures compared to P2P systems, which typically include message signing,
hash linking, and extra verification rules (as evaluated in
Section~\ref{sec.evaluation}).

@@

Regarding P2P protocols, Bitcoin~\cite{p2p.bitcoin} is probably the most
successful permissionless network, but serves specifically for electronic
cash.
IPFS~\cite{p2p.ipfs} and Dat~\cite{p2p.dat} are data-centric protocols for
hosting large files and applications, respectively.
Scuttlebutt~\cite{p2p.scuttlebutt} and Aether~\cite{p2p.ecosystem} are closer
to \FC goals and cover human communication.

Bitcoin adopts proof-of-work to achieve consensus, which does not solve the
centralization issue entirely, given the high costs of equipment and energy.
Proof-of-stake is a prominent alternative~\cite{p2p.proofs} that acknowledges
that centralization is inevitable, and thus uses a function of time and wealth
to elect peers to mint new blocks.
As an advantage, these proof mechanisms are generic and apply to multiple
domains, since they depend on extrinsic resources.
% (i.e., the richer gets richer)
In contrast, we chose an intrinsic resource, which is authored content in the
chains themselves.
We believe that human work grows more linearly with effort and is not directly
portable across chains with different topics.
These hypotheses support the intended decentralization of our system.
%
Another distinction is that generic public ledgers require permanent
connectivity to avoid forks, which opposes our local-first principle.
This is because a token transaction only has value as part of the longest
chain.
This is not the case for a local message exchange between friends, which has
value in itself.
